---
title: Practical Machine Learning Coursera Course Project
subtitle    : Predicting the manner in which people do excercise 
author: Diego Alonso Lazo Paz
job: Stutdent
output: pdf_document
framework   : io2012        
highlighter : highlight.js  
hitheme     : tomorrow       
url:
  lib: ../../librariesNew
  assets: ../../assets
widgets     : [mathjax]            
mode        : selfcontained 
---
```{r,include=FALSE}
library(caret);library(dplyr);library(ggplot2);library(gbm);library(rpart);library(rpart.plot);library(rattle);library(RColorBrewer);library(corrplot);library(explore);library(DataExplorer);library(naniar);library(MLmetrics)
```

```{r,echo=TRUE}
train<-read.csv("pml-training.csv")
test<-read.csv("pml-testing.csv")
val_df<-test
```

## Background
* Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.
* The objective of this project is to predict the manner in which people do excercise using different machine learning models. 

## Partition of the traning data
```{r,fig.width=6.5, fig.height=3,echo=TRUE}
indx  <- createDataPartition(train$classe, p=0.7, list=FALSE)
Traindf0 <- train[indx, ]
Testdf0  <- train[-indx, ]
```


## Exploratory data Analysis
* In this part we look to understand the data in order to obtain insights and have a first look into the behaviour of the variables
```{r,fig.width=6.5, fig.height=3,echo=TRUE}
Traindf0 %>% explore_tbl()
introduce(Traindf0)
```
* We 19.6k observations and  160 variables, which 123 are double or integer, the rest are characters.
* 61 double variables have NA values.
```{r,fig.width=6.5, fig.height=3, echo=TRUE}
plot_intro(Traindf0)
```
* So we are going to remove the variables with too many missing values and also the the variables with near zero variance
```{r, echo=TRUE}
#Remove variable with many NAS
Traindf1<- Testdf0[, colSums(is.na(Testdf0)) == 0]
Testdf1 <- Testdf0[, colSums(is.na(Testdf0)) == 0]
# remove variables with Nearly Zero Variance
near_zero_var <- nearZeroVar(Traindf1)
Traindf2 <- Traindf1[, -near_zero_var]
Testdf2  <- Testdf1[, -near_zero_var]
# remove identification only variables (columns 1 to 5)
Traindf3 <- Traindf2[, -(1:5)]
Testdf3  <- Testdf2[, -(1:5)]
```

### Correlation Analysis
```{r,fig.width=6.5, fig.height=3, echo=TRUE}
corMatrix <- cor(Traindf3[, -54])
corrplot(corMatrix, order = "FPC", method = "color", type = "lower", 
         tl.cex = 0.8, tl.col = rgb(0, 0, 0))
```
* The plot is not much of a help , so with use a function to find the variables with the highest correlation

```{r,fig.width=6.5, fig.height=3, echo=TRUE}
highlyCorrelated = findCorrelation(corMatrix, cutoff=0.80)
names(Traindf3)[highlyCorrelated]
```

## Predictive Modelling
* In order to obtain the best prediction, we are going to use differente models to get the best prediction, and if it is possible, combine predictions to get the best prediction possible
* For this part we are going to use the next models: simple classification tree (rpart), K-Nearest Neighbords (knn), random forest and a Generalized Boosted Regression Models(gbm)

### Classification tree
```{r,fig.width=6.5, fig.height=3, echo=TRUE}
set.seed(123)
class_tree <- rpart(classe ~ ., data=Traindf3, method="class")
fancyRpartPlot(class_tree)
```

```{r,fig.width=6.5, fig.height=3, echo=TRUE}
predict_clstree <- predict(class_tree, Testdf3, type = "class")
conf_mtree <- confusionMatrix(predict_clstree, Testdf3$classe %>% as.factor())
conf_mtree
plot(conf_mtree$table, col = conf_mtree$byClass, main = paste("Accuracy =", round(conf_mtree$overall['Accuracy'], 4)))
```
* With this simple model we get an accuracy of 0.82

### K-Nearest Neighbords
```{r, echo=TRUE}
set.seed(12345)
ctrlKNN <- trainControl(method="cv", number=5, classProbs= TRUE, summaryFunction = multiClassSummary)
m_kknn <- train(classe~., 
                data=Traindf3,
                trControl = ctrlKNN,
                method="kknn" )
m_kknn$finalModel
predict_knn <- predict(m_kknn, newdata=Testdf3)
cmkn <- confusionMatrix(predict_knn, Testdf3$classe %>% as.factor())
cmkn
```

### Random Forest
```{r,fig.width=6.5, fig.height=3, echo=TRUE}
set.seed(12345)
ctrlRF <- trainControl(method="cv", number=5, verboseIter=FALSE)
rf_model <- train(classe ~ ., 
                  data=Traindf3, 
                  trControl = ctrlRF,
                  method="rf")
rf_model$finalModel
predict_RF <- predict(rf_model, newdata=Testdf3)
cmrf <- confusionMatrix(predict_RF, Testdf3$classe %>% as.factor())
cmrf
plot(rf_model)

```

### Generalized Boosted Regression Models(gbm)
```{r,fig.width=6.5, fig.height=3, echo=TRUE}
set.seed(12345)
ctrlGBM <- trainControl(method = "repeatedcv", number = 5)
gbm_model <- train(classe ~ ., 
                  data=Traindf3, 
                  trControl = ctrlGBM,
                  method="gbm",
                  verbose=FALSE)
gbm_model$finalModel
predict_gbm <- predict(gbm_model, newdata=Testdf3)
cmgbm <- confusionMatrix(predict_gbm, Testdf3$classe %>% as.factor())
cmgbm
plot(gbm_model)
```

## Apply the best model to validation data
The accuracy of the 4 classification models  are:
Decision Tree : 0.8054
KNN: 0.999
Random Forest : 0.999
GBM : 0.9932
* We choose the GBM because de RF and KNN are overfitting the data

```{r,fig.width=6.5, fig.height=3, echo=TRUE}
Results <- predict(gbm_model, newdata=val_df)
Results
```

